---
title: "Untitled"
output: html_document
---

```{r load package}

library(tidyverse)

```


#Behavioural Data

```{r get data}

#Instant and cont behaviours grouped by day of observation
#comes from DataPreparation_Scan file
Scan <- Scan_Session_Data_Day %>% 
  rename(AnimalID = Subject)

#Point and state by session 
Focal_SessionSummary <- 
  
Focal_DaySummary <- 
  
  

```


##Scan 

```{r Subset}

#BS experimental data 

#Are there any experimental data we may want remove

```

##Focal 


#Individual Info
I wonder whether that should be also part of a pipeline. The data prep of each project would then simply be to subset the data

##Static
```{r List}

############################################Scan 
#Animal Observed during scan
Subject_Scan <- Scan %>% 
  distinct(AnimalID)
View(List_Subject_Scan)
#1127 animals observed

############################################Focal


############################################Merge Scan & Focal

```

```{r Static}
Static <- idinfo_static(Subject_Scan)
```

## Dynamic

Dynamic info I will be interested in 
- Age 
- Weight 
- Breeding status: animal considered as breeder if already gave birth in the lab.
- Gestation status: could be done at a later stage as it will depend on analyses

```{r List}

############################################Scan 
#Observed Animal/Date combination during scan
SubjectDate_Scan <- Scan %>% 
  distinct(AnimalID,ObsDate)
View(List_SubjectDate_Scan)
#18989 combinations

#Observed Animal/Date combination during scan + static info
#Will serve to get dynamic info
SubjectDateStatic_Scan <- List_SubjectDate_Scan %>% 
  left_join(.,Static)

############################################Focal 


############################################Merge

```

Age, breeding status and weight can be queried quite easily from the static information. They could wither be piped in Static or queried separately with stand alone function. 
Gestation and lactation status are more difficult and may require group characteristics
```{r Dynamic}

#######################################Weight
#Closest weight
#1 weight for animal/date
ClosestWeight <- SubjectDate_Scan %>% 
  closestweight(Date = ObsDate)
View(ClosestWeight)

#######################################Age
#Age
#1 age per animal/date
Age <- SubjectDateStatic_Scan %>% 
  select(AnimalID,ObsDate,BirthDate) %>% 
  add_age(date = ObsDate, unit = "months") %>% 
  add_age(date = ObsDate, unit = "days")
View(Age)


#######################################Breeding status 
#Breeding Status
#1 row per animal/date
BreedingStatus <- SubjectDateStatic_Scan %>% 
  select(AnimalID,ObsDate,Sex,FirstConceptionDate) %>% 
  add_breedingstatus(date = ObsDate, firstconceptiondate = FirstConceptionDate)
View(BreedingStatus)

View(BreedingStatus %>% 
       filter(Sex == "F",
              BreedingStatus == ))


#######################################Next Parturition
#Delay til Next parturition with dates
#1 row for each breeding female/date when ObsDate > FirstConception AND ObsDate < Parturition
NextParturition <- 
  #FUNCTION DATA
  #RETAIN FEMALE THAT BRED
  SubjectDateStatic_Scan %>%
  filter(Sex == "F",
         !(is.na(FirstConceptionDate))) %>%
  select(AnimalID,ObsDate,FirstParturitionDate) %>% 
  #NEXT PARTURITION ()
  next_parturitioninfo(., date = ObsDate) %>% 
  #SELECT 
  select(AnimalID,ObsDate,Next_ParturitionDate,Next_ParturitionDayDiff)
View(NextParturition)
#1660
#2618

#######################################GestationStatus
#at the moment the function does not deal with the undetermined period

#Semester
GestationStatus_Semester <-   SubjectDateStatic_Scan %>%
  #FUNCTION DATA
  #RETAIN FEMALE THAT BRED
  filter(Sex == "F",
         !(is.na(FirstConceptionDate))) %>% 
  select(AnimalID, ObsDate, DeathDate, FirstParturitionDate, LastParturitionDate) %>% 
  #ADD DAYDIFF TO NEXT PARTURITION
  left_join(.,NextParturition) %>% 
  mutate(BreedingTenure_EndDate = case_when(!(is.na(DeathDate)) ~ DeathDate, 
                                              TRUE ~ today())) %>%
  gestationstatus(date = ObsDate,gestationsplit = "semester") %>% 
  arrange(AnimalID)
View(GestationStatus_Semester)

#Trimester
GestationStatus_Trimester <-   SubjectDateStatic_Scan %>%
  #FUNCTION DATA
  #RETAIN FEMALE THAT BRED
  filter(Sex == "F",
         !(is.na(FirstConceptionDate))) %>% 
  select(AnimalID, ObsDate, DeathDate, FirstParturitionDate, LastParturitionDate) %>% 
  #ADD DAYDIFF TO NEXT PARTURITION
  left_join(.,NextParturition) %>% 
  mutate(BreedingTenure_EndDate = case_when(!(is.na(DeathDate)) ~ DeathDate, 
                                              TRUE ~ today())) %>%
  gestationstatus(date = ObsDate,gestationsplit = "trimester") %>% 
  arrange(AnimalID)
View(GestationStatus_Trimester)
 


View(GestationStatus)
   
  
  
  
  
  
    #ADD STARTDATE UNDETERMINED PERIOD 
    mutate(Gestation_Undetermined_StartDate = case_when(BreedingTenure_EndDate - 90 >= LastParturitionDate ~ BreedingTenure_EndDate - 90,
           BreedingTenure_EndDate - 90 < LastParturitionDate ~ LastParturitionDate)) %>% 
  
          mutate(GestationStatus_Trimester = case_when(Next_ParturitionDayDiff < -90  ~ "NotPregnant", 
                                           Next_ParturitionDayDiff >= - 90 & Next_ParturitionDayDiff < -(90/2) ~ "FirstSemester",
                                           Next_ParturitionDayDiff >= -(90/2) & Next_ParturitionDayDiff <=0 ~ "SecondSemester",
                                           is.na(Next_ParturitionDayDiff) & ObsDate >= Gestation_Undetermined_StartDate ~"Undetermined"))

View(poo)
#2810
  




List_SubjectDateStatic_Scan %>% 
  select(AnimalID,ObsDate,FirstConceptionDate)mutate(BreedingTenure_EndDate = case_when(!(is.na(DeathDate)) ~ DeathDate, 
                                              TRUE ~ today())) %>% 
    #ADD STARTDATE UNDETERMINED PERIOD 
    mutate(Gestation_Un





########################################Join all dynamic data

# #all dynamic info
# Dynamic <- List_SubjectDate_Scan %>% 
#   #ADD STATIC
#   left_join(.,Static) %>% 
#   select(AnimalID,ObsDate,BirthDate,FirstConceptionDate,Queen) %>% 
#   #ADD AGE
#   add_age(date = ObsDate, unit = "months") %>% 
#   add_age(date = ObsDate, unit = "days") %>% 
#   #ADD WEIGHT
#   left_join(.,ClosestWeight) %>% 
#   #ADD BREEDING STATUS
#   add_breedingstatus(date = ObsDate)
#   
# 
# 
#   left_join(.,gestationstatus(DF = AnimalID_Static_FocalDates)) %>% 
# 
# View(Dynamic)

```

#Group Info
- Group size 
- Pup presence 
- Pup number 
- Pup age 
- Breeding status of group (through data prep)

```{r list}

#List of group and dates
List_GroupDate_Scan <- Scan %>% 
  distinct(Colony,ObsDate)
View(List_GroupDate_Scan)
#2324

```

```{r Group comp}

#Get group comp 
GroupComp_Scan <- groupcomp(List_GroupDate_Scan, Date = ObsDate)
View(GroupComp_Scan)
```

```{r Static}

#Static info of group comp 
GroupComp_Static_Scan <- GroupComp_Scan %>% 
  left_join(.,Animal_Info)
View(GroupComp_Static_Scan)
#27'03
```

```{r Dynamic}

#Age

#Weight 

#BreedingStatus

```

```{r Group}

#####################################Group Size 
#GroupSize
GroupSize <- groupsize(GroupComp_Scan, Date = ObsDate)
View(GroupSize_Scan)
#2'319 loss of 5 rows

#Number of pups 
PupNumber <- 

#Age of pups 

#Number of pups 

#breeding status 

```

