---
title: "Untitled"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r package}

rm(list = ls())

#Data tidying
library(tidyverse)


#Data
library(broom)
library(tidyverse)
library(cowplot)#
library(ggcorrplot)#
library(PerformanceAnalytics)#For correlation plot
library(gridExtra)#
library(FactoMineR)#
library(factoextra)#

```


#Info and functions
```{r data and functions}

#extract relevant info from database
#don't run the new behaviour import if not necessary because it takes for ages
source("C:/Users/Philippe/Dropbox/PhD_20150515/Research projects/DMR_Ratop/Database_Extract_20210113.R")

#Load relevant functions
source("C:/Users/Philippe/Dropbox/PhD_20150515/Research projects/DMR_Ratop/DMRfunctions_20210202.R")

#Prepare most relevant data
#I could already have all
source("C:/Users/Philippe/Dropbox/PhD_20150515/Research projects/DMR_Ratop/DataPreparation_20200113.R")

```


#Data preparation

Once a good pipeline is in place we may be able to move that to data preparation and functions

##Scan

Old and new data are not merged yet. Until it is done, it makes no sense to focus on continuous data as I guess that would be a loss of time

```{r Scan data}

#scan session details
#Old scan session details are available from Scan_SessionDetails, which comes from user_philippev
View(Scan_SessionDetails)
names(Scan_SessionDetails)
#1650 scan sessions

#Get behaviours from all scan observations performed so far
#Continuous behaviours are not analyses friendly and something should be done about it
Scan_Behaviours <- scan_behaviour_count(Scan_SessionDetails)
View(Scan_Behaviours)


#Join Session details and behaviours
Scan_ColonyBehaviour <- Scan_SessionDetails %>% 
  select(Colony,ObsRef,ObsDate) %>% 
  left_join(.,Scan_Behaviours)
View(Scan_ColonyBehaviour)



#Continuous behaviours
# I will have to generate the zero
ScanCont_Behaviours <- Scan_Behaviours %>% 
  filter(ObsType == "ScanCont") %>% 
  distinct(Behaviour)
View(ScanCont_Behaviours)
#Continuous behaviours during scan old import consist of 
#Biting                 AnimalID            BodyPart
#Carry Pup              PupID
#Chase                  AnimalID
#Coprophagy             Never recorded
#Copulation             AnimalID            Success           Direction
#DropFS                 
#Food comp              AnimalID            Location          FoodStore
#Groom.                 AnimalID            Duration          Reciprocation
#Nest Build             Carry/Build
#Overt Aggress          AnimalID
#Pass                   AnimalID            Direction         Over/Under
#Pull Tail              AnimalID
#Pump.                  Number
#Sex Forplay            AnimalID            Duration
#Sniff.                 AnimalID            Location          BodyPart
#Sparr                  AnimalID            Duration
#Submiss                AnimalID            Count             Context

#I think pass, copulation, Nest build? could be divided by their divider 

#the earliest dropFS comes from 15th June 2016, not sure how consistent it was at the beginning


```


##Animal Info
```{r Animal Info }

#######################################List of Colony/AnimalID/Date

#Queried group composition on days of observation
#It doesnt mean that all animals queried were part of scan
#it is also possible that some animals may be wrongly assigned to a given colony and were not observed
Scan_GroupComp_Queried <- groupcomp(Scan_SessionDetails %>% 
  select(ObsRef,Colony,ObsDate), Date = ObsDate) %>% 
    distinct(ObsDate,Colony,AnimalID)
View(Scan_GroupComp_Queried)
#13537


#Get group composition based on which animals were observed during a scan session 
#It could be that the composition is not exact as some animals may have been excluded from the scan observation
Scan_GroupComp_Observed <- Scan_SessionDetails %>% 
  select(ObsRef,Colony,ObsDate) %>% 
  left_join(.,Scan_Behaviours %>%
              select(ObsRef,AnimalID)) %>% 
  distinct(ObsDate,Colony,AnimalID) %>% 
  mutate(AnimalID = toupper(AnimalID))
View(Scan_GroupComp_Observed)
#13'199


#Get group composition after merging group comp queried and animal observed
GroupComp_Scan <- bind_rows(Scan_GroupComp_Queried,Scan_GroupComp_Observed) %>% 
  distinct()
View(GroupComp_Scan)
#13'572


#######################################Static info

#fixed characteristics of individual 
AnimalID_Static_Scan <- idinfo_static(GroupComp_Scan %>% 
  distinct(AnimalID))
View(AnimalID_Static_Scan)
#in the old data 890 individuals have been observed 


#######################################Dynamic info

#Get a list of dates of interest, AnimalID and static info 
#Should have 13'572 rows
AnimalID_Static_ScanDates<- GroupComp_Scan %>%
  left_join(.,AnimalID_Static_Scan)
View(AnimalID_Static_ScanDates)
#13567, loss of 5 rows


#Individual Info
#Age <- age_inmonths()
#Breeding Status <- breedingstatus 
#Gestation status <- gestationstatus(); function is not terminated to determine breeding status after last parturition
#weight <- clsosestweight()
AnimalID_Dynamic_Scan <- AnimalID_Static_ScanDates %>% 
  age_inmonths(Date = ObsDate) %>% 
  breedingstatus(Date = ObsDate) %>% 
  left_join(.,gestationstatus(DF = AnimalID_Static_ScanDates)) %>% 
  left_join(.,closestweight(DF1 = AnimalID_Static_ScanDates %>% 
                              select(AnimalID,ObsDate), DF2 = tblWeights, Date = ObsDate))
View(AnimalID_Dynamic_Scan)
#13'567


#Ask Colin why it doesn't work
# poo <-  AnimalID_Scan_Dynamic %>% 
#     mutate(EndDate = case_when(!(is.na(DeathDate)) ~ DeathDate,
#                                TRUE ~ today())) %>% #assign an "End Date for each female
#     left_join(.,Litter_Info %>% 
#                  filter(Sex == "F") %>% 
#                  select(AnimalID, IndividualConceptionCount_Total, OffspringLitterRef, ParturitionDate)) %>% #shall I add conception or parturition group or whatever?
#   mutate(ParturitionCollection_DayDiff = as.integer(ObsDate - ParturitionDate),#delay between date of interest and parturition
#          ) %>% 
#   filter(ObsDate >= FirstParturitionDate -105) %>% #one wants 
#   #GROUP BY ANIMALID, DATE
#   group_by(AnimalID, ObsDate) %>% 
#   filter(ObsDate < ParturitionDate & ParturitionCollection_DayDiff == max(ParturitionCollection_DayDiff)) #Why doesn't it work?
```


## Group Info 

```{r Group Info}

#Get info on group conposition
#A current limitation is that all individuals who were wild caught are considered older than 18 months which is not going to be true for quite a lot of them.
GroupCompInfo_Scan <- groupinfo(DF = AnimalID_Dynamic_Scan %>% 
  select(ObsDate,Colony,AnimalID,Sex,Age_Months), Date = ObsDate, Age = Age_Months)
View(GroupCompInfo_Scan)

#Breeding status of group
#is actually necessary to get the correct gestation status (see in function)


#One thing that will surely be relevant is the size of the tunnel system. One will have to look with Tom where he had that information
#This will surely need an update from Dave Seager. Have they recorded the information? If yes where? 
```


##All Info

```{r All Info}

AllInfo_Scan <- AnimalID_Dynamic_Scan %>% 
  left_join(.,GroupCompInfo_Scan) %>% 
  mutate(Lactating = case_when(Sex == "F" & BreedingStatus ==  "Breeder" & Pup14 > 0 ~ "Lactating", 
                               TRUE ~ "NotLactating"))

```


#Analyses

We will mostly be interested in determining whether behavioural differences are mostly driven by gestation and lactation.

The main questions we will be asking are: 
Q1: what are the behavioural differences between BF and NBF; Activity? Dominance? Aggressive behaviour? Huddling pups?
Q2: Are behavioural differences magnified by gestation, lactation

Data visualization:
V1: violin plot for scan instantaneous
V2: frequency for scan continuous
V3: Violin for focal state
V4: 

Analyses ideas: 
A1: Correlations between all behaviours (scan and focal) using PCA, heatmat etc
A2: Multinomial logistic regression to analyse Scan data instantaneous 
A3: Are BF more aggressive than NBF
A4: Are BF dominant over NBF 
A5: effects of pups/lactation on activity 


##Scan
Since scan have not been merged yet, I will focus on instantaneous sampling. I will take care of continuous sampling once merging will have been done

###DataPrep

```{r Scan Data Prep}

#Select the animal info we are likely going to be interested in
SelectedInfo_Scan <- AllInfo %>% 
  select(ObsDate,Colony,AnimalID,WildCaught,Sex,Age_Months,Weight,BreedingStatus,GestationStatus,Lactating)
View(SelectedInfo_Scan)


#########################################Behaviours by session

#17 behaviours 
#As it stands total scan number includes pup carry
#if it is removed, total scan number should be recalculated
ScanInst_BySession <- Scan_ColonyBehaviour %>% 
  filter(ObsType == "ScanInst",
         !(Behaviour %in% c("Help.Burrow","Active.NoHelp","Help.Total","TotActive"))) %>% 
  arrange(ObsRef,AnimalID,Behaviour) %>% 
  mutate(Behaviour = case_when(Behaviour == "Food.Carry" ~ "Foodcarry",
                               Behaviour == "Nest.Material" ~ "Nestbuilding",
                               Behaviour == "Pup.Carry" ~ "Pupcarry",
                               Behaviour == "Self.Groom" ~ "Selfgroom",
                               Behaviour == "Social.Interact" ~ "Socint",
                               TRUE ~ Behaviour)) %>% 
  mutate(Proportion = Count/TotalScanNumber,
         Category7 = case_when(Behaviour %in% c("Rest","Huddeling") ~ "Rest",
                               Behaviour %in% c("Dig","Sweep","Kick","LocoWork") ~ "Burrow",
                               Behaviour %in% c("Gnaw","Other","Pump","Selfgroom","Sniff","Locomotion","Socint") ~ "Activeother",
                               TRUE ~ Behaviour)) %>% 
  select(-c(ObsType,Received)) %>% 
  arrange(Colony,ObsRef,AnimalID,Category7,Behaviour)
View(ScanInst_BySession)


#Join Info of relevance with scan behaviours 
#animals that will have no behaviours will have been supposed to be in groups but were not observed
ScanInst_BySession_WithInfo <- SelectedInfo_Scan %>% 
  left_join(.,ScanInst_BySession) %>% 
  relocate(ObsRef,.after = ObsDate)
View(ScanInst_BySession_WithInfo)


#Count by categories
#This could be used for the mulytinomial model
ScanInst7_BySession_WithInfo <- ScanInst_BySession_WithInfo %>% 
  #GROUP BY OBSREF, ANIMALID, BEHAVIOUR CATEGORY
  group_by(ObsRef, AnimalID, Category7) %>% 
  mutate(Count_Category = sum(Count)) %>% 
  ungroup() %>% 
  #UNGROUP
  select(-Count,-Behaviour, -Proportion) %>% 
  distinct() %>% 
  mutate(Proportion_Category = Count_Category/TotalScanNumber)
View(ScanInst7_BySession_WithInfo)



#########################################Data check/Exploration

#pupcarry
View(ScanInst_BySession %>% 
       filter(Behaviour == "Pupcarry", 
              Proportion > 0))
#There is only 98 entries of pupcarry out of 233'376 entries. It is totally negligible


#animals in group comp but for which there is no behaviour, suggesting behaviours may not have imported and/or these animals were excluded from scan
View(ScanInst_BySession_WithInfo %>% 
       filter(is.na(ObsRef))) 
View(ScanInst_BySession_WithInfo %>% 
       filter(is.na(ObsRef))%>% 
       distinct(ObsDate,Colony,AnimalID))
#374 animals that are missing behaviours
#sometimes it looks like it is entire scan others just a few animals

```


###Data Visualisation 

Violin plot: with 17 behaviours, with 7 categories
Heat map
Correlation plot

```{r}

######################################Violin plot

######################################Heat map correlation structure
#Not sure what it will bring 
#An idea could be to enforce clustering by categories and see whether behavioural pattern differes from when clustering is not enforced by categories (disregarding categories)

######################################Correlation plot 

```


###PCA

###LDA 

###Multinomial 


