---
title: "Untitled"
output: html_document
editor_options: 
  chunk_output_type: console
---


```{r package}

rm(list = ls())

#Data tidying
library(tidyverse)

```


#Info and functions
```{r data and functions}

#extract relevant info from database
#don't run the new behaviour import if not necessary because it takes for ages
source("C:/Users/Philippe/Dropbox/PhD_20150515/Research projects/DMR_Ratop/Database_Extract_20210305.R")

#Load relevant functions
source("C:/Users/Philippe/Dropbox/PhD_20150515/Research projects/DMR_Ratop/DMRfunctions_20210305.R")

#Prepare most relevant data
source("C:/Users/Philippe/Dropbox/PhD_20150515/Research projects/DMR_Ratop/DataPreparation_20210305.R")

```

#Data preparation

##Animal Info
Query all informations of animals that were observed within the frame of focal observations. A better alternative may be to query information on entire group compositions if I end up needing info on group comp for example. This will most likely depend from type of analyses

```{r Focal ID List}

####################################Possibly observed animals
#based on group composition queried from colony and date entered as independent variable qry_FocalSession_All

#Observation date 
Focal_ColonyDate <- qry_FocalSession_All %>% 
  distinct(Colony,FocalDate)

#Query group composition using groupID recorded as independent variable in Observer
#Do I need to add MemberFrom and MemberTo here?
Focal_GroupComp_Queried <- groupcomp(DF1 = Focal_ColonyDate, Date = FocalDate)
View(Focal_GroupComp_Queried)


####################################Effectively Observed animals
#Colony is the one from qry_FocalSession_All

#Focal ID
Focal_FocalID <- FocalRaw_Dyads_Cleaned %>% 
  distinct(FocalDate,Colony,AnimalID) %>% 
  filter(!(is.na(AnimalID))) %>% 
  mutate(AnimalType = "Focal") %>%
  arrange(AnimalID,FocalDate)
View(Focal_FocalID)
#6251

#Partner ID
Focal_PartnerID <- FocalRaw_Dyads_Cleaned %>% 
  distinct(FocalDate,Colony,PartnerID) %>% 
  rename(AnimalID = PartnerID) %>% 
  mutate(AnimalType = "Partner") 
View(Focal_PartnerID)
#18349

#All Observed animals 
#Merge FocalID and PartnerID
Focal_AnimalList_Observed <-  bind_rows(Focal_FocalID,Focal_PartnerID) %>%
  select(-AnimalType) %>% 
  distinct()
View(Focal_AnimalList_Observed)
#20'774 unique combination of AnimalID and dates

```

Get static information on each animal that was observed in the course of focal observations. At this stage, I have only done it for observed animals.

```{r Static}
Focal_ObservedAnimals_Static <- Focal_AnimalList_Observed %>%
  distinct(AnimalID) %>% 
  idinfo_static()
View(Focal_ObservedAnimals_Static)
#926 animals were observed

```

Get dynamic information on each animal that was observed in the course of focal observations
```{r Dynamic}
#Needs to start with a list of AnimalID and dates

#Get a list of dates of interest, AnimalID and static info 
#Should have 13'572 rows
Focal_ObservedAnimals_Dates <- Focal_AnimalList_Observed %>%
  select(AnimalID,FocalDate) %>% 
  left_join(.,Focal_ObservedAnimals_Static)
View(Focal_ObservedAnimals_Dates)
#20'774 no loss of rows


#Dynamic info
#Age <- age_inmonths()
#Breeding Status <- breedingstatus 
#Gestation status <- gestationstatus() breeding status after last parturition
#weight <- closestweight()
Focal_AnimalID_Dynamic <- Focal_ObservedAnimals_Dates %>%
  age_inmonths(Date = FocalDate) %>% 
  breedingstatus(Date = FocalDate) %>% 
  left_join(.,gestationstatus(DF = Focal_ObservedAnimals_Dates,Date = FocalDate)) %>% 
  distinct() %>% ##from 20'774 to 20'784 (no distinct) and 20'758 if distinct. Not sure why it creates a duplication 
  left_join(.,closestweight(DF1 = Focal_ObservedAnimals_Dates %>% 
                              select(AnimalID,FocalDate), DF2 = tblWeights, Date = FocalDate))
```


##Group info
At this stage I am not sure if and what group info may be needed. It will depend on analyses. For example do ERGM need individual info on each individual at a given date of observation or not? 


##Behaviours

The membership of Partner ID is needed to exclude interactions that may have been impossible. 

For each behaviour, I will have to generate a count of interaction they have been part of. I must have previously done it with Colin. Go and find the code on how it was done.

```{r Focal Membership}

####################################Queried membership
#Membership will return double entry if date correspond to a date of change

#All Observed animals membership
Focal_FocalPartnerID_Membership <- Focal_AnimalList_Observed %>% 
  select(AnimalID,FocalDate,AnimalType) %>% 
  membership(Date = FocalDate) %>% 
  rename(QueriedColony = Colony)
View(Focal_FocalPartnerID_Membership)


#FocalID queried membership
#includes HBO2 animals
Focal_FocalID_Membership <- Focal_FocalID %>% 
  select(AnimalID,FocalDate,AnimalType) %>% 
  membership(Date = FocalDate) %>% 
  rename(QueriedColony = Colony)
View(Focal_FocalID_Membership)


#PartnerID queried membership  
Focal_PartnerID_Membership <- Focal_PartnerID %>% 
  select(AnimalID,FocalDate,AnimalType) %>% 
  membership(Date = FocalDate) %>% 
  rename(QueriedColony = Colony)
View(Focal_PartnerID_Membership)

```

```{r Dyads}

#FocalRaw_Dyads_Cleaned from data prep file
#Remove interactions where partnerID has a queried colony that differ from Colony entered as an independent variable in FocalSession
Dyads <- FocalRaw_Dyads_Cleaned %>%
  filter(ExpName != "HBO2") %>% #For network analyses these can be excluded
  inner_join(.,Focal_FocalPartnerID_Membership %>% 
              select(AnimalID,FocalDate,QueriedColony),by=c("PartnerID" = "AnimalID","FocalDate" = "FocalDate")) %>% 
  distinct() %>% 
  filter(Colony == QueriedColony)#Only retain interactions where colony as independent variable is the same is Queried colony of Partner ID 

```


###Submissive calls

```{r data prep}
#select sub calls
#shape for further analyses
SubCall <- Dyads %>% 
  filter(Behaviour == "Submiss call") %>% 
  select(DataSource,FocalRef,FocalDate,StartTime,EventTime,Behaviour,Colony,AnimalID,PartnerID,ReceivedValue,InteractionRef)

#received submissive call
SubCall_R <- SubCall %>% 
  filter(ReceivedValue == 0) %>% 
  rename(WinnerID = AnimalID,
         LoserID = PartnerID) %>% 
  select(-ReceivedValue)

#Given submissive call
SubCall_G <- SubCall %>% 
  filter(ReceivedValue == 1) %>% 
  rename(WinnerID = PartnerID,
         LoserID = AnimalID) %>% 
  select(-ReceivedValue)

#Merge received and given value 
SubCall_EloDF <- bind_rows(SubCall_R,SubCall_G) %>% 
  arrange(Colony,FocalDate,StartTime,EventTime)#check whether that is going to give me the same sequence
View(SubCall_EloDF)

write.csv(SubCall_EloDF,"SubCall_EloDF_20210803.csv",row.names = FALSE)
```


###Pass
So far I have excluded sidewards pass but they could be used as null outcome
```{r data prep}

#select sub calls
#shape for further analyses
Pass_Front <- Dyads %>% 
  filter(Behaviour == "Pass",
         ModifierLabel2 == "NoseToNose") %>% 
  select(DataSource,FocalRef,FocalDate,StartTime,EventTime,Behaviour,Colony,AnimalID,PartnerID,ModifierLabel3,InteractionRef)
View(Pass_Front)

#received submissive call
PassOver <- Pass_Front %>% 
  filter(ModifierLabel3 == "Over") %>% 
  rename(WinnerID = AnimalID,
         LoserID = PartnerID)
View(PassOver)

#Given submissive call
PassUnder <- Pass_Front %>% 
  filter(ModifierLabel3 == "Under") %>% 
  rename(WinnerID = PartnerID,
         LoserID = AnimalID)
View(PassUnder)

#Merge received and given value 
Pass_EloDF <- bind_rows(PassOver,PassUnder) %>% 
  arrange(Colony,FocalDate,StartTime,EventTime)#check whether that is going to give me the same sequence
View(Pass_EloDF)

write.csv(SubCall_EloDF,"Pass_EloDF_20210803.csv",row.names = FALSE)

```

#Analyses 
I will have to join behaviour info on animal info. Based on analyses, I will use winner/loser or Focal/Partner to join. Since I am not sure yet, I have not done it

##Elo calculation

###Sub Call 


###Pass